---
title: "Bluefish in the NMFS bottom trawl"
author: "Abigail Tyrell"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

devtools::load_all()
devtools::load_all(here::here("../bluefishLifeHistory"))
`%>%` <- magrittr::`%>%`

```

## Tow date, latitude, longitude, and surface temperature vary by year
```{r}
dat <- bluefish_nmfs %>% 
  dplyr::mutate(DOY = lubridate::yday(EST_TOWDATE),
                size_group = ifelse(LENGTH <= 32.1,
                                    "small (<=32.1cm)",
                                    ifelse(LENGTH >= 49.6,
                                           "large (>=49.6cm)",
                                           "medium (31.2-49.6cm)"))) %>%
  tidyr::drop_na(SURFTEMP) %>% 
  dplyr::filter(SURFTEMP > 0,
                SEASON == "FALL")#,
                #DOY < 305) # don't include November

dat %>% 
  ggplot2::ggplot(ggplot2::aes(x = DOY, 
                               y = SURFTEMP, 
                               color = as.factor(YEAR))) + 
  ggplot2::geom_point() + 
  ggplot2::geom_smooth(method = "lm", se = FALSE) +
  ggplot2::facet_wrap(~size_group, scales = "free")

dat %>% 
  ggplot2::ggplot(ggplot2::aes(x = DOY, 
                               y = LAT, 
                               color = as.factor(YEAR))) + 
  ggplot2::geom_point() + 
  ggplot2::geom_smooth(method = "lm", se = FALSE) +
  ggplot2::facet_wrap(~size_group, scales = "free")

dat %>% 
  ggplot2::ggplot(ggplot2::aes(x = DOY, 
                               y = LON, 
                               color = as.factor(YEAR))) + 
  ggplot2::geom_point() + 
  ggplot2::geom_smooth(method = "lm", se = FALSE) +
  ggplot2::facet_wrap(~size_group, scales = "free")
```

```{r}
library(mgcv)
correct_temp <- function(data, formula, gam_formula) {

  print("glm")
  model <- gam(as.formula(formula), 
               data = data,
               method = "REML")
  par(mfrow = c(2,2))
 # plot(model)
  par(mfrow = c(1,1))
  pred <- ggeffects::ggpredict(model, 
                                terms = "YEAR",
                                condition = c(NUMLEN = 1))
  print(pred)

  cat("\n\n")
  # plot(pred) %>% print()
  # cat("\n\n")
  
  print("gam")
  model2 <- gam(as.formula(gam_formula), 
            data = data,
            method = "REML")
  gam.check(model2)
    par(mfrow = c(2,2))
  plot(model2, all.terms = TRUE, rug = TRUE) %>%
    try()
  par(mfrow = c(1,1))
  pred2 <- ggeffects::ggpredict(model2, 
                                terms = "YEAR",
                                condition = c(NUMLEN = 1))
  print(pred2)
  cat("\n\n")
  # plot(pred2) %>% print()
  # cat("\n\n")
  
  compare_data <- rbind(pred %>% dplyr::mutate(type = "GLM"),
                        pred2 %>% dplyr::mutate(type = "GAM"))
  
  plt <- compare_data %>%
    ggplot2::ggplot(ggplot2::aes(x = x %>%
                                   as.character() %>%
                                   as.numeric(),
                                 y = predicted,
                                 color = type,
                                 group = type)) +
    ggplot2::geom_line() +
    ggplot2::geom_point() +
    ggplot2::theme_bw()
  
  print(plt)
  
  # plot(pred$x, pred$predicted, type = "b")
  # points(pred2$x, pred2$predicted, col = "red")
  # lines(pred2$x, pred2$predicted, col = "red")
  
  AIC(model, model2)
}

dat <- bluefish_nmfs %>%
  tidyr::drop_na(LENGTH, SURFTEMP) %>%
  dplyr::mutate(size_group = ifelse(LENGTH <= 32.1,
                                    "small (<=32.1cm)",
                                    ifelse(LENGTH >= 49.6,
                                           "large (>=49.6cm)",
                                           "medium (31.2-49.6cm)")),
                DOY = lubridate::yday(EST_TOWDATE)) %>%
  tidyr::drop_na(SURFTEMP) %>%
  dplyr::filter(SURFTEMP > 0,
                SEASON == "FALL") %>%
  dplyr::group_by(EST_TOWDATE,YEAR, SEASON, size_group, LAT, LON) %>%
  dplyr::summarise(n_fish = sum(NUMLEN)) %>%
  dplyr::ungroup()

## check for normality, outliers

par(mfrow = c(1,2))

boxplot(dat$DOY)
hist(dat$DOY)

boxplot(dat$LAT)
hist(dat$LAT)

boxplot(dat$LON)
hist(dat$LON)

boxplot(dat$SURFTEMP)
hist(dat$SURFTEMP)

par(mfrow = c(1,1))

# remove outlier LAT > 44
# dat <- dat %>%
#   dplyr::filter(LAT < 44)

# # add zeros
all_tows <- all_nmfs %>%
  tidyr::drop_na(SURFTEMP) %>%
  dplyr::filter(SEASON == "FALL",
                SURFTEMP > 0) %>% 
  dplyr::select(SEASON, EST_TOWDATE, LAT, LON, SURFTEMP, YEAR) %>%
  dplyr::mutate(size_group = "no fish",
                n_fish = 0,
                DOY = lubridate::yday(EST_TOWDATE))

zeros <- dplyr::anti_join(all_tows, dat, by = "EST_TOWDATE") 

full_bluefish <- dplyr::full_join(zeros, dat) %>%
  dplyr::select(SEASON, EST_TOWDATE, LAT, LON, SURFTEMP, YEAR, size_group, n_fish, DOY) %>%
  dplyr::group_by(YEAR, SEASON, size_group) %>%
  dplyr::mutate(annual_group_fish = sum(n_fish),
                annual_group_event = dplyr::n()) %>%
  dplyr::ungroup() %>%
   dplyr::group_by(YEAR) %>%
   dplyr::mutate(lat_range = max(LAT) - min(LAT),
                 annual_all_fish = sum(annual_group_fish),
                 annual_all_event = sum(annual_group_event)) %>%
   dplyr::filter(lat_range > 6,
                 annual_all_fish > 0)

hurdle_model <- gam(list(n_fish ~ s(LAT) + s(LON) + te(LAT, LON, k = 7) + DOY + SURFTEMP*as.factor(YEAR) + size_group,
          ~ DOY + SURFTEMP*as.factor(YEAR)),
          data = full_bluefish,
          family = ziplss)
summary(hurdle_model)



# model_dat <- dat %>%
#                    dplyr::filter(size_group == "small (<=32.1cm)") %>%
#   dplyr::full_join(zeros) %>%
#   dplyr::select(SURFTEMP, LAT, LON, YEAR, DOY, NUMLEN)

# correct_temp(data = model_dat,
#              formula = "SURFTEMP ~ LAT*LON + DOY + NUMLEN*as.factor(YEAR)",
#              gam_formula = "SURFTEMP ~ s(LAT) + s(LON) + te(LAT, LON, k = 7) + DOY + NUMLEN*as.factor(YEAR)")

# correct_temp(data = model_dat,
#              formula = "NUMLEN ~ LAT*LON + SURFTEMP*DOY + as.factor(YEAR)*DOY + NUMLEN",
#              gam_formula = "NUMLEN ~  s(LAT) + s(LON) + te(LAT, LON, k = 7) + s(SURFTEMP) + te(SURFTEMP, DOY, k = 7) + as.factor(YEAR)*DOY")

small <- dat %>% 
  dplyr::filter(size_group == "small (<=32.1cm)") %>%
  dplyr::full_join(zeros) %>%
  dplyr::select(SURFTEMP, LAT, LON, YEAR, DOY, NUMLEN) %>%
   dplyr::group_by(YEAR) %>%
   dplyr::mutate(lat_range = max(LAT) - min(LAT),
                 n_fish = sum(NUMLEN)) %>%
   dplyr::filter(lat_range > 6,
                 n_fish > 0)
medium <- dat %>% 
  dplyr::filter(size_group == "medium (31.2-49.6cm)") %>%
  dplyr::full_join(zeros) %>%
  dplyr::select(SURFTEMP, LAT, LON, YEAR, DOY, NUMLEN) %>%
   dplyr::group_by(YEAR) %>%
   dplyr::mutate(lat_range = max(LAT) - min(LAT),
                 n_fish = sum(NUMLEN)) %>%
   dplyr::filter(lat_range > 6,
                 n_fish > 0)
large <- dat %>% 
  dplyr::filter(size_group == "large (>=49.6cm)") %>%
  dplyr::full_join(zeros) %>%
  dplyr::select(SURFTEMP, LAT, LON, YEAR, DOY, NUMLEN) %>%
   dplyr::group_by(YEAR) %>%
   dplyr::mutate(lat_range = max(LAT) - min(LAT),
                 n_fish = sum(NUMLEN)) %>%
   dplyr::filter(lat_range > 6,
                 n_fish > 0)
```

## Surface temperature

Try to account for annual variation in latitude, longitude, and day of year when looking at how surface temperature preferences change over time

### Small bluefish
```{r}
correct_temp(data = small,
             formula = "SURFTEMP ~ LAT*LON + DOY + NUMLEN*as.factor(YEAR)",
             gam_formula = "SURFTEMP ~ s(LAT) + s(LON) + te(LAT, LON, k = 7) + DOY + NUMLEN*as.factor(YEAR)")
```

### Medium bluefish
```{r}
correct_temp(data = medium,
             formula = "SURFTEMP ~ LAT*LON + DOY + NUMLEN*as.factor(YEAR)",
             gam_formula = "SURFTEMP ~ s(LAT) + s(LON) + te(LAT, LON, k = 7) + DOY + NUMLEN*as.factor(YEAR)")
```

### Large bluefish
```{r}
correct_temp(data = large,
             formula = "SURFTEMP ~ LAT*LON + DOY + NUMLEN*as.factor(YEAR)",
             gam_formula = "SURFTEMP ~ s(LAT) + s(LON) + te(LAT, LON, k = 7) + DOY + NUMLEN*as.factor(YEAR)")
```

## Latitude

Try to account for annual variation in longitude and day of year when looking at how latitude preferences change over time

### Small bluefish
```{r}
correct_temp(data = small,
             formula = "LAT ~ LON + NUMLEN*as.factor(YEAR)",
             gam_formula = "LAT ~ s(LON) + NUMLEN*as.factor(YEAR)")
```

### Medium bluefish
```{r}
correct_temp(data = medium,
             formula = "LAT ~ LON + NUMLEN*as.factor(YEAR)",
             gam_formula = "LAT ~ s(LON) + NUMLEN*as.factor(YEAR)")
```

### Large bluefish
```{r}
correct_temp(data = large,
             formula = "LAT ~ LON + NUMLEN*as.factor(YEAR)",
             gam_formula = "LAT ~ s(LON) + NUMLEN*as.factor(YEAR)")
```

## Longitude

Try to account for annual variation in latitude and day of year when looking at how longitude preferences change over time

### Small bluefish
```{r}
correct_temp(data = small,
             formula = "LON ~ LAT + as.factor(YEAR)*NUMLEN",
             gam_formula = "LON ~ s(LAT) + as.factor(YEAR)*NUMLEN")
```

### Medium bluefish
```{r}
correct_temp(data = medium,
             formula = "LON ~ LAT + as.factor(YEAR)*NUMLEN",
             gam_formula = "LON ~ s(LAT) + as.factor(YEAR)*NUMLEN")
```

### Large bluefish
```{r}
correct_temp(data = large,
             formula = "LON ~ LAT + as.factor(YEAR)*NUMLEN",
             gam_formula = "LON ~ s(LAT) + as.factor(YEAR)*NUMLEN")
```

## ESP figures

### Small bluefish

```{r, echo = FALSE, fig.height = 3, message = FALSE}
## small fish
small %>%
  gam(formula = SURFTEMP ~ s(LAT) + s(LON) + te(LAT, LON, k = 7) + DOY + NUMLEN*as.factor(YEAR), 
      family = gaussian()) %>%
  ggeffects::ggpredict(terms = "YEAR", 
                       condition = c(NUMLEN = 1)) %>%
  tibble::as_tibble() %>%
  dplyr::select(x, predicted) %>%
  dplyr::rename(YEAR = x,
                DATA_VALUE = predicted) %>%
  dplyr::mutate(INDICATOR_NAME = "fall temperature, small (<=31.2cm)",
                YEAR = YEAR %>%
                  as.character() %>%
                  as.numeric()) %>%
  bluefishESP::esp_traffic_fig(status = FALSE,
                               out = "one_pager")
```

```{r, echo = FALSE, fig.height = 3, message = FALSE}
small %>%
  gam(formula = LAT ~ s(LON) + as.factor(YEAR)*NUMLEN, 
      family = gaussian()) %>%
  ggeffects::ggpredict(terms = "YEAR", 
                       condition = c(NUMLEN = 1)) %>%
  tibble::as_tibble() %>%
  dplyr::select(x, predicted) %>%
  dplyr::rename(YEAR = x,
                DATA_VALUE = predicted) %>%
  dplyr::mutate(INDICATOR_NAME = "fall latitude, small (<=31.2cm)",
                YEAR = YEAR %>%
                  as.character() %>%
                  as.numeric()) %>%
  bluefishESP::esp_traffic_fig(status = FALSE,
                               out = "one_pager")
```

```{r, echo = FALSE, fig.height = 3, message = FALSE}
small %>%
  gam(formula = LON ~ s(LAT) + as.factor(YEAR)*NUMLEN, 
      family = gaussian()) %>%
  ggeffects::ggpredict(terms = "YEAR", 
                       condition = c(NUMLEN = 1)) %>%
  tibble::as_tibble() %>%
  dplyr::select(x, predicted) %>%
  dplyr::rename(YEAR = x,
                DATA_VALUE = predicted) %>%
  dplyr::mutate(INDICATOR_NAME = "fall longitude, small (<=31.2cm)",
                YEAR = YEAR %>%
                  as.character() %>%
                  as.numeric()) %>%
  bluefishESP::esp_traffic_fig(status = FALSE,
                               out = "one_pager")
```

### Medium bluefish
```{r, echo = FALSE, fig.height = 3, message = FALSE}
medium %>%
  gam(formula = SURFTEMP ~ s(LAT) + s(LON) + te(LAT, LON, k = 7) + DOY + NUMLEN*as.factor(YEAR), 
      family = gaussian()) %>%
  ggeffects::ggpredict(terms = "YEAR", 
                       condition = c(NUMLEN = 1)) %>%
  tibble::as_tibble() %>%
  dplyr::select(x, predicted) %>%
  dplyr::rename(YEAR = x,
                DATA_VALUE = predicted) %>%
  dplyr::mutate(INDICATOR_NAME = "fall temperature, medium (31.2-49.6cm)",
                YEAR = YEAR %>%
                  as.character() %>%
                  as.numeric()) %>%
  bluefishESP::esp_traffic_fig(status = FALSE,
                               out = "one_pager")
```

```{r, echo = FALSE, fig.height = 3, message = FALSE}
medium %>%
  gam(formula = LAT ~ s(LON) + as.factor(YEAR)*NUMLEN, 
      family = gaussian()) %>%
  ggeffects::ggpredict(terms = "YEAR", 
                       condition = c(NUMLEN = 1)) %>%
  tibble::as_tibble() %>%
  dplyr::select(x, predicted) %>%
  dplyr::rename(YEAR = x,
                DATA_VALUE = predicted) %>%
  dplyr::mutate(INDICATOR_NAME = "fall latitude, medium (31.2-49.6cm)",
                YEAR = YEAR %>%
                  as.character() %>%
                  as.numeric()) %>%
  bluefishESP::esp_traffic_fig(status = FALSE,
                               out = "one_pager")
```

```{r, echo = FALSE, fig.height = 3, message = FALSE}
medium %>%
  gam(formula = LON ~ s(LAT) + as.factor(YEAR)*NUMLEN, 
      family = gaussian()) %>%
  ggeffects::ggpredict(terms = "YEAR", 
                       condition = c(NUMLEN = 1)) %>%
  tibble::as_tibble() %>%
  dplyr::select(x, predicted) %>%
  dplyr::rename(YEAR = x,
                DATA_VALUE = predicted) %>%
  dplyr::mutate(INDICATOR_NAME = "fall longitude, medium (31.2-49.6cm)",
                YEAR = YEAR %>%
                  as.character() %>%
                  as.numeric()) %>%
  bluefishESP::esp_traffic_fig(status = FALSE,
                               out = "one_pager")
```

### Large bluefish
```{r, echo = FALSE, fig.height = 3, message = FALSE}
## large fish
large %>%
  gam(formula = SURFTEMP ~ s(LAT) + s(LON) + te(LAT, LON, k = 7) + DOY + NUMLEN*as.factor(YEAR), 
      family = gaussian()) %>%
  ggeffects::ggpredict(terms = "YEAR", 
                       condition = c(NUMLEN = 1)) %>%
  tibble::as_tibble() %>%
  dplyr::select(x, predicted) %>%
  dplyr::rename(YEAR = x,
                DATA_VALUE = predicted) %>%
  dplyr::mutate(INDICATOR_NAME = "fall temperature, large (>=49.6cm)",
                YEAR = YEAR %>%
                  as.character() %>%
                  as.numeric()) %>%
  bluefishESP::esp_traffic_fig(status = FALSE,
                               out = "one_pager")
```

```{r, echo = FALSE, fig.height = 3, message = FALSE}
large %>%
  gam(formula = LAT ~ s(LON) + as.factor(YEAR)*NUMLEN, 
      family = gaussian()) %>%
  ggeffects::ggpredict(terms = "YEAR", 
                       condition = c(NUMLEN = 1)) %>%
  tibble::as_tibble() %>%
  dplyr::select(x, predicted) %>%
  dplyr::rename(YEAR = x,
                DATA_VALUE = predicted) %>%
  dplyr::mutate(INDICATOR_NAME = "fall latitude, large (>=49.6cm)",
                YEAR = YEAR %>%
                  as.character() %>%
                  as.numeric()) %>%
  bluefishESP::esp_traffic_fig(status = FALSE,
                               out = "one_pager")
```

```{r, echo = FALSE, fig.height = 3, message = FALSE}
large %>%
  gam(formula = LON ~ s(LAT) + as.factor(YEAR)*NUMLEN, 
      family = gaussian()) %>%
  ggeffects::ggpredict(terms = "YEAR", 
                       condition = c(NUMLEN = 1)) %>%
  tibble::as_tibble() %>%
  dplyr::select(x, predicted) %>%
  dplyr::rename(YEAR = x,
                DATA_VALUE = predicted) %>%
  dplyr::mutate(INDICATOR_NAME = "fall longitude, large (>=49.6cm)",
                YEAR = YEAR %>%
                  as.character() %>%
                  as.numeric()) %>%
  bluefishESP::esp_traffic_fig(status = FALSE,
                               out = "one_pager")
```
