# Things that have been removed to save space

# Additional figures


## Annual state catches

```{r, eval = TRUE}
state_rec_catch <- params$mrip_catch %>%
  dplyr::group_by(year, st_f) %>%
  dplyr::summarise(DATA_VALUE = sum(tot_cat, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
  dplyr::mutate(name = paste(st_f, "_recreational_catch_n", sep = " \n ")) %>%
  dplyr::rename(YEAR = year) %>%
  dplyr::group_by(name) %>%
  dplyr::mutate(mean = mean(DATA_VALUE, na.rm = TRUE),
                sd = sd (DATA_VALUE, na.rm = TRUE),
                YEAR = as.numeric(YEAR)) %>%
  dplyr::ungroup()

state_rec_catch$name <- factor(state_rec_catch$name,
                               levels = paste(c("MAINE", "NEW HAMPSHIRE",
                                                "MASSACHUSETTS", "RHODE ISLAND",
                                                "CONNECTICUT", "NEW YORK",
                                                "NEW JERSEY", "DELAWARE",
                                                "MARYLAND", "VIRGINIA",
                                                "NORTH CAROLINA", "SOUTH CAROLINA",
                                                "GEORGIA", "FLORIDA"), 
                                              "_recreational_catch_n", 
                                              sep = " \n "))
```

#### Northeast 
```{r}
state_rec_catch %>%
  dplyr::filter(st_f %in% c("MAINE", "NEW HAMPSHIRE",
                                                "MASSACHUSETTS", "RHODE ISLAND",
                                                "CONNECTICUT")) %>%
  dplyr::rename(INDICATOR_NAME = name) %>%
  esp_traffic_fig(status = FALSE) %>% 
  try()
```

#### Mid Atlantic
```{r}
state_rec_catch %>%
  dplyr::filter(st_f %in% c("NEW YORK",
                            "NEW JERSEY", "DELAWARE",
                            "MARYLAND", "VIRGINIA")) %>%
  dplyr::rename(INDICATOR_NAME = name) %>%
  esp_traffic_fig(status = FALSE) %>% 
  try()
```

#### Southeast
```{r}
state_rec_catch %>%
  dplyr::filter(st_f %in% c("NORTH CAROLINA", "SOUTH CAROLINA",
                            "GEORGIA", "FLORIDA")) %>%
  dplyr::rename(INDICATOR_NAME = name) %>%
  esp_traffic_fig(status = FALSE) %>% 
  try()
```


```{r, eval = FALSE}
plt1 <- data %>%
  dplyr::filter(wave_f != "JAN/FEB") %>%
  ggplot2::ggplot(ggplot2::aes(x = st_f,
                               y = mean_prop_coast_catch,
                               fill = mean_state_catch)) +
  ggplot2::geom_col(color = "black") +
  ggplot2::coord_flip() +
  ggplot2::facet_wrap(~wave_f,
                      ncol = 1
           ) +
  ggplot2::theme_bw() +
  viridis::scale_fill_viridis(name = "Mean fish caught\n(number)") +
  ggplot2::xlab(ggplot2::element_blank()) +
  ggplot2::ylab("Mean proportion of coastwide catch") +
  ggplot2::theme(aspect.ratio = 0.5,
                 axis.text.y = ggplot2::element_text(size = 8))

plt2 <- data %>%
  dplyr::filter(wave_f != "JAN/FEB") %>%
  ggplot2::ggplot(ggplot2::aes(x = st_f,
                               y = unstandardized_CPUE,
                               fill = mean_state_catch)) +
  ggplot2::geom_col(color = "black") +
  ggplot2::coord_flip() +
  ggplot2::facet_wrap(~wave_f,
                      ncol = 1
           ) +
  ggplot2::theme_bw() +
  viridis::scale_fill_viridis(name = "Mean fish caught\n(number)") +
  ggplot2::xlab(ggplot2::element_blank()) +
  ggplot2::ylab("Unstandardized CPUE") +
  ggplot2::theme(axis.text.y = ggplot2::element_blank(),
                 aspect.ratio = 0.5)

plt3 <- data %>%
  dplyr::filter(wave_f != "JAN/FEB") %>%
  ggplot2::ggplot(ggplot2::aes(x = st_f,
                               y = standardized_CPUE,
                               fill = mean_state_catch)) +
   ggplot2::geom_col(ggplot2::aes(y = ),
                     color = "black") +
  ggplot2::coord_flip() +
  ggplot2::facet_wrap(~wave_f,
                      ncol = 1
           ) +
  ggplot2::theme_bw() +
  viridis::scale_fill_viridis(name = "Mean fish caught\n(number)") +
  ggplot2::xlab(ggplot2::element_blank()) +
  ggplot2::ylab("Standardized CPUE") +
  ggplot2::theme(axis.text.y = ggplot2::element_blank(),
                 aspect.ratio = 0.5)

ggpubr::ggarrange(plt1, plt2, plt3, 
                  ncol = 3,
                  common.legend = TRUE,
                  legend = "bottom",
                  widths = c(1.1, 1, 1),
                  heights = 1)
```

### Commercial correlations with bluefish stock
```{r, fig.height = 3}
bluefish <- params$data %>%
  dplyr::filter(INDICATOR_NAME == "Commercial_Bluefish_Landings_LBS") %>%
  dplyr::select(YEAR, DATA_VALUE) %>%
  dplyr::rename(Comm_landings = DATA_VALUE)

fuel <- params$data %>%
  dplyr::filter(INDICATOR_NAME == "AVG_ANNUAL_DIESEL_FUEL_PRICE_REAL_DOLLARS") %>%
  dplyr::select(YEAR, DATA_VALUE) %>%
  dplyr::rename(Fuel_price = DATA_VALUE)

dplyr::full_join(bluefish, fuel, by = "YEAR") %>%
  tidyr::drop_na(Comm_landings, Fuel_price) %>%
  plot_corr(x = "Fuel_price",
          y = "Comm_landings",
          title = "Commercial landings vs fuel")
```

### Recreational correlations with bluefish stock
```{r, fig.height = 3}
# catch
params$data %>%
  dplyr::filter(INDICATOR_NAME == "total_recreational_catch_n") %>%
  dplyr::select(YEAR, DATA_VALUE) %>% 
    dplyr::rename(Rec_catch = DATA_VALUE) %>%
  dplyr::full_join(fuel, by = "YEAR") %>%
  tidyr::drop_na(Rec_catch, Fuel_price) %>%
  plot_corr(x = "Fuel_price",
          y = "Rec_catch",
          title = "Recreational catch vs fuel")
cat("\n\n")

# landings
params$data %>%
  dplyr::filter(INDICATOR_NAME == "total_recreational_landings_lbs") %>%
  dplyr::select(YEAR, DATA_VALUE) %>%
  dplyr::rename(Rec_landings = DATA_VALUE) %>%
  dplyr::full_join(fuel, by = "YEAR") %>%
  tidyr::drop_na(Rec_landings, Fuel_price) %>%
  plot_corr(x = "Fuel_price",
          y = "Rec_landings",
          title = "Recreational landings vs fuel")
cat("\n\n")

# trips
params$data %>%
  dplyr::filter(INDICATOR_NAME == "Bluefish_trips") %>%
  dplyr::select(YEAR, DATA_VALUE) %>%
  dplyr::rename(Rec_trips = DATA_VALUE) %>%
  dplyr::full_join(fuel, by = "YEAR") %>%
  tidyr::drop_na(Rec_trips, Fuel_price) %>%
  plot_corr(x = "Fuel_price",
          y = "Rec_trips",
          title = "Recreational trips vs fuel")
cat("\n\n")
```

### Climate correlations with bluefish stock

#### Correlation plots
```{r, fig.height = 3}
bluefish <- NEesp::asmt %>%
  dplyr::filter(Species == "Bluefish",
                Metric == "Recruitment" |
                  Metric == "Biomass",
                AssessmentYear == 2019) %>%
  dplyr::select(Year, Metric, Value) %>%
  tidyr::pivot_wider(names_from = "Metric", values_from = "Value") %>%
  dplyr::mutate(rssb = Recruitment/Biomass) %>%
  dplyr::rename(YEAR = Year) %>%
  dplyr::select(YEAR, rssb)

for(i in c("first_18c_day", "last_18c_day", "n_18c_day", "july_proportion_22c",
           "crossshore_wind", "longshore_wind"
)){
  params$data %>%
  dplyr::filter(INDICATOR_NAME == i) %>%
  dplyr::full_join(bluefish, by = "YEAR") %>%
  tidyr::drop_na(rssb, DATA_VALUE) %>%
  plot_corr(x = "DATA_VALUE",
          y = "rssb",
          title = paste("Bluefish Recruitment per SSB (metric ton) vs", i)) %>%
    print()
cat("\n\n")
           }
```

#### GLM
```{r, eval = FALSE}
ssb <- NEesp::asmt %>%
  dplyr::filter(Species == "Bluefish",
                Metric == "Recruitment",
                AssessmentYear == 2019) %>%
  dplyr::rename(Bluefish_SSB = Value,
                YEAR = Year) %>%
  dplyr::select(YEAR, Bluefish_SSB)

climate_data <- params$data %>%
  dplyr::filter(INDICATOR_NAME %in% c("first_18_day", "last_18_day", "n_18_day", "july_proportion",
           "crossshore_wind_April", "crossshore_wind_May", 
           "longshore_wind_April", "longshore_wind_May"
))  %>%
  tidyr::pivot_wider(names_from = "INDICATOR_NAME", values_from = "DATA_VALUE") %>%
  dplyr::full_join(bluefish) %>%
  dplyr::full_join(ssb) %>%
  tidyr::drop_na()

mod0 <- glm(Bluefish_Recruitment ~ 1,
            data = climate_data,
            family = "poisson")
mod0b <- glmmTMB::glmmTMB(Bluefish_Recruitment ~ 1,
            data = climate_data,
            family = glmmTMB::nbinom1())

# MASS::stepAIC(mod0,
#               scope = list(upper = ~ crossshore_wind_April_North + crossshore_wind_May_North + crossshore_wind_April_South + crossshore_wind_May_South + longshore_wind_April_North + longshore_wind_May_North + longshore_wind_April_South + longshore_wind_May_South + first_18_day + last_18_day + n_18_day + july_proportion + Bluefish_SSB),
#               direction = "forward",
#               trace = -1) %>%
#   invisible()

mod <- glm(Bluefish_Recruitment ~ crossshore_wind_April + crossshore_wind_May + + longshore_wind_April + longshore_wind_May + first_18c_day + last_18c_day + n_18c_day + july_propotion_22C + Bluefish_SSB,
            data = climate_data,
            family = "poisson")
modb <- glmmTMB::glmmTMB(Bluefish_Recruitment ~ crossshore_wind_April + crossshore_wind_May + + longshore_wind_April + longshore_wind_May + first_18c_day + last_18c_day + n_18c_day + july_propotion_22C + Bluefish_SSB,
            data = climate_data,
            family = glmmTMB::nbinom1())

AIC(mod0, mod0b, mod, modb)
```

```{r, results = "markup", eval = FALSE}
summary(mod)
```

```{r, fig.height = 6, eval = FALSE}
plot(climate_data$YEAR, climate_data$Bluefish_Recruitment, type = "b")

mod_prediction <- exp(predict(mod))

lines(climate_data$YEAR, mod_prediction, col = "red")
```

Probably has too many degrees of freedom

### Mortality correlations with blufish stock
```{r, fig.height = 3}
# mako
bluefish <- NEesp::asmt %>%
  dplyr::filter(Species == "Bluefish",
                Metric == "Biomass",
                AssessmentYear == 2019) %>%
  dplyr::rename(DATA_VALUE = Value,
                YEAR = Year) %>%
  dplyr::select(YEAR, DATA_VALUE) %>%
  dplyr::rename(Bluefish_SSB = DATA_VALUE)

params$data %>%
  dplyr::filter(INDICATOR_NAME == "mako_avg_bbmsy") %>%
  dplyr::full_join(bluefish, by = "YEAR") %>%
  tidyr::drop_na(Bluefish_SSB, DATA_VALUE) %>%
    dplyr::rename(Mako_BBmsy = DATA_VALUE) %>%
  plot_corr(x = "Mako_BBmsy",
          y = "Bluefish_SSB",
          title = "Bluefish SSB vs Mako shark B/Bmsy")

# condition
for(i in c(
  # condition indicator names
)){
  params$data %>%
  dplyr::filter(INDICATOR_NAME == i) %>%
  dplyr::full_join(bluefish, by = "YEAR") %>%
  tidyr::drop_na(Bluefish_Recruitment, DATA_VALUE) %>%
  plot_corr(x = "DATA_VALUE",
          y = "Bluefish_SSB",
          title = paste("Bluefish SSB vs", i)) %>%
    print()
cat("\n\n")
}
```

## NEAMAP geography
#### Histogram
```{r, fig.height = 12, eval = FALSE}
neamap <- read.csv(here::here("data-raw/dont-share/State Data Sets 2021/")) %>%
    dplyr::mutate(size_group = ifelse(is.na(LENGTH), "unknown",
                                      ifelse(LENGTH <= 35, "small (<=35cm)", "large (>35cm)")),
                  size_class = floor(LENGTH/10) * 10,
                month = format(EST_TOWDATE, "%m"),
                Decade = floor(YEAR/10) * 10) 

fig <- nmfs %>%
  ggplot2::ggplot(ggplot2::aes(x = LAT,
                               fill = as.factor(size_class)
                              # fill = size_group
                               )) +
  ggplot2::geom_histogram(color = "black",
                          binwidth = 1#,
                    # position = ggplot2::position_dodge()
                     ) +
  ggplot2::facet_wrap(~month,
                      ncol = 2,
                      #rows = ggplot2::vars(month),
                      #cols = ggplot2::vars(Decade),
                      scales = "free_y") +
  ggplot2::theme_bw()

print(fig)
```

#### Map
```{r, fig.height = 20, fig.width = 8, eval = FALSE}
ggplot2::ggplot(nmfs %>% 
                  dplyr::group_by(YEAR, month, size_group, LAT, LON) %>%
                  dplyr::summarise(n_fish = sum(NUMLEN)),
                       ggplot2::aes(
                         x = LON,
                         y = LAT)) +
  ggplot2::geom_point(ggplot2::aes(size = n_fish,
                      fill = as.factor(size_group)),
                      pch = 21) +
  ggplot2::geom_sf(data = rnaturalearth::ne_countries(scale = "medium", 
                                                      returnclass = "sf"),
                   inherit.aes = FALSE) +
  ggplot2::coord_sf(
    xlim = c(min(nmfs$LON) - 1, max(nmfs$LON) + 1),
    ylim = c(min(nmfs$LAT) - 1, max(nmfs$LAT) + 1)
  ) +
  ggplot2::xlab("Longitude") +
  ggplot2::ylab("Latitude") +
  ggplot2::facet_grid(cols = ggplot2::vars(size_group),
                      rows = ggplot2::vars(month)) +
  ggplot2::theme_bw() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90),
                 legend.position = "top")
```


## NEFSC bottom trawl

### Bluefish latitude
20 days only
```{r, fig.height = 4}
params$data %>%
  dplyr::filter(stringr::str_detect(INDICATOR_NAME, "lat")) %>%
  dplyr::mutate(SEASON = ifelse(stringr::str_detect(INDICATOR_NAME, "FALL"),
                                 "FALL", "SPRING"),
                size_group = ifelse(stringr::str_detect(INDICATOR_NAME, ">35cm"),
                                 "large (>35cm)", "small (<=35cm)"),
                statistic = ifelse(stringr::str_detect(INDICATOR_NAME, "max"),
                                 "max",
                              ifelse(stringr::str_detect(INDICATOR_NAME, "min"),
                                 "min", "mean"))
                ) %>%
  ggplot2::ggplot(ggplot2::aes(x = YEAR,
                               y = DATA_VALUE,
                               color = statistic)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(method = "lm", se = FALSE
  ) +
  ggplot2::facet_grid(cols = ggplot2::vars(size_group), 
                      rows = ggplot2::vars(paste(SEASON))) +
  ggplot2::theme_bw()
```

### Bluefish temperature
20 days only
```{r, fig.height = 4}
params$data %>%
  dplyr::filter(stringr::str_detect(INDICATOR_NAME, "temp")) %>%
  dplyr::mutate(SEASON = ifelse(stringr::str_detect(INDICATOR_NAME, "FALL"),
                                 "FALL", "SPRING"),
                size_group = ifelse(stringr::str_detect(INDICATOR_NAME, ">35cm"),
                                 "large (>35cm)", "small (<=35cm)"),
                statistic = ifelse(stringr::str_detect(INDICATOR_NAME, "max"),
                                 "max",
                              ifelse(stringr::str_detect(INDICATOR_NAME, "min"),
                                 "min", "mean"))
                ) %>%
  ggplot2::ggplot(ggplot2::aes(x = YEAR,
                               y = DATA_VALUE,
                               color = statistic)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(method = "lm", se = FALSE
  ) +
  ggplot2::facet_grid(cols = ggplot2::vars(size_group), 
                      rows = ggplot2::vars(paste(SEASON))) +
  ggplot2::theme_bw() 
```

### Weekly data
All weeks with more than 9 samples
```{r, warning = FALSE}
dat <- params$bluefish_nmfs %>%
  tidyr::drop_na(LENGTH) %>%
  dplyr::mutate(size_class = trunc(LENGTH/10) * 10,
                size_group = ifelse(LENGTH <= 35, "small (<=35cm)", "large (>35cm)"),
                month = format(EST_TOWDATE, "%m"),
                week = lubridate::week(EST_TOWDATE)) %>%
  dplyr::group_by(YEAR, week, size_group) %>%
  dplyr::summarise(min_lat = min(LAT, na.rm = TRUE),
                   max_lat = max(LAT, na.rm = TRUE),
                   n_sample_events = sum(!is.na(LAT)),
                   weighted_mean_lat = mean(rep(LAT, NUMLEN), na.rm = TRUE),
                   min_temp = min(SURFTEMP, na.rm = TRUE),
                   max_temp = max(SURFTEMP, na.rm = TRUE),
                   weighted_mean_temp = mean(rep(SURFTEMP, NUMLEN), na.rm = TRUE)) %>%
  dplyr::filter(n_sample_events > 5) %>%
  tidyr::pivot_longer(cols = c("min_lat", "max_lat", "weighted_mean_lat",
                               "min_temp", "max_temp", "weighted_mean_temp")) %>%
  dplyr::mutate(INDICATOR_NAME = paste(week, size_group, name)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(INDICATOR_NAME) %>%
  dplyr::filter(dplyr::n() > 9)


fig <- dat %>%
    dplyr::filter(stringr::str_detect(name, "lat")) %>%
  ggplot2::ggplot(ggplot2::aes(x = YEAR,
                               y = value,
                               color = name)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(method = "lm", se = FALSE
  ) +
  ggplot2::facet_grid(cols = ggplot2::vars(size_group), 
                      rows = ggplot2::vars(week)) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "top")

  print(fig)
  
    fig <- dat %>%
    dplyr::filter(stringr::str_detect(name, "temp")) %>%
  ggplot2::ggplot(ggplot2::aes(x = YEAR,
                               y = value,
                               color = name)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(method = "lm", se = FALSE
  ) +
  ggplot2::facet_grid(cols = ggplot2::vars(size_group), 
                      rows = ggplot2::vars(week)) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "top")

  print(fig)
```

#### Histogram
```{r, fig.height = 10}
nmfs <- params$bluefish_nmfs %>%
    dplyr::mutate(size_group = ifelse(is.na(LENGTH), "unknown",
                                      ifelse(LENGTH <= 35, "small (<=35cm)", "large (>35cm)")),
                  size_class = floor(LENGTH/10) * 10,
                month = format(EST_TOWDATE, "%m"),
                Decade = floor(YEAR/10) * 10) 

fig <- nmfs %>%
  ggplot2::ggplot(ggplot2::aes(x = LAT,
                               fill = as.factor(size_group)
                              # fill = size_group
                               )) +
  ggplot2::geom_histogram(color = "black",
                          binwidth = 1#,
                    # position = ggplot2::position_dodge()
                     ) +
  ggplot2::facet_wrap(~month,
                      ncol = 2,
                      #rows = ggplot2::vars(month),
                      #cols = ggplot2::vars(Decade),
                      scales = "free_y") +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "top")

print(fig)
```

#### Map
```{r, fig.height = 20, fig.width = 8}
ggplot2::ggplot(nmfs %>% 
                  dplyr::group_by(YEAR, month, size_group, LAT, LON) %>%
                  dplyr::summarise(n_fish = sum(NUMLEN)),
                       ggplot2::aes(
                         x = LON,
                         y = LAT)) +
  ggplot2::geom_point(ggplot2::aes(size = n_fish,
                      fill = as.factor(size_group)),
                      pch = 21) +
  ggplot2::geom_sf(data = rnaturalearth::ne_countries(scale = "medium", 
                                                      returnclass = "sf"),
                   inherit.aes = FALSE) +
  ggplot2::coord_sf(
    xlim = c(min(nmfs$LON) - 1, max(nmfs$LON) + 1),
    ylim = c(min(nmfs$LAT) - 1, max(nmfs$LAT) + 1)
  ) +
  ggplot2::xlab("Longitude") +
  ggplot2::ylab("Latitude") +
  ggplot2::facet_grid(cols = ggplot2::vars(size_group),
                      rows = ggplot2::vars(month)) +
  ggplot2::theme_bw() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90),
                 legend.position = "top")
```

## Recreational catch

#### By wave
```{r}
state_wave <- params$mrip_catch %>%
  dplyr::group_by(year, st_f, wave_f) %>%
  dplyr::summarise(DATA_VALUE = sum(tot_cat, na.rm = TRUE)) %>%
  dplyr::mutate(name = paste0(st_f, "\nrecreational_catch_n\n", wave_f)) %>%
  dplyr::rename(YEAR = year) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(name) %>%
  dplyr::mutate(mean = mean(DATA_VALUE, na.rm = TRUE),
                sd = sd (DATA_VALUE, na.rm = TRUE),
                YEAR = as.numeric(YEAR))

for(i in c("JAN/FEB", "MAR/APR", "MAY/JUN", "JUL/AUG", "SEP/OCT", "NOV/DEC")){
  message(i)
  
  this_data <- state_wave %>%
    dplyr::filter(wave_f == i,
                  !st_f %in% c("ALABAMA", "LOUISIANA", "MISSISSIPPI")) %>%
    dplyr::ungroup() %>%
    dplyr::group_by(st_f) %>%
    dplyr::mutate(n_years = dplyr::n()) %>%
    dplyr::filter(n_years > 2)
  
  
  this_data$name <- factor(this_data$name,
                               levels = paste0(c("MAINE", "NEW HAMPSHIRE",
                                                "MASSACHUSETTS", "RHODE ISLAND",
                                                "CONNECTICUT", "NEW YORK",
                                                "NEW JERSEY", "DELAWARE",
                                                "MARYLAND", "VIRGINIA",
                                                "NORTH CAROLINA", "SOUTH CAROLINA",
                                                "GEORGIA", "FLORIDA"), 
                                               "\nrecreational_catch_n\n",
                                               i))
 fig <- this_data %>%
   dplyr::rename(INDICATOR_NAME = name) %>%
  esp_traffic_fig(paginate = TRUE,
                  status = FALSE,
                  ncolumn = 2)
  
  print(fig)
}
```

## Landings per trip 

#### By wave
```{r}
# set up data
cpue <- dplyr::full_join(wave_catch,
                         wave_trips,
                         by = c("wave_f", "st_f", "year")) %>%
  dplyr::group_by(st_f, year, wave_f) %>%
  dplyr::summarise(wave_catch = sum(state_catch_n, na.rm = TRUE),
                   wave_land = sum(state_land_lbs, na.rm = TRUE),
                   wave_trips = sum(trips, na.rm = TRUE)) %>%
  dplyr::mutate(n_per_trip = wave_catch/wave_trips,
                lbs_per_trip = wave_land/wave_trips) %>%
  dplyr::select(-c(wave_catch, wave_land, wave_trips)) %>%
  tidyr::pivot_longer(cols = c("n_per_trip", "lbs_per_trip")) %>%
  dplyr::mutate(name = paste(st_f, wave_f, name, sep = " \n ")) %>%
  dplyr::rename(YEAR = year, 
                DATA_VALUE = value) %>%
  dplyr::ungroup() # %>%
  # dplyr::group_by(name) %>%
  # dplyr::mutate(mean = mean(DATA_VALUE, na.rm = TRUE),
  #               sd = sd (DATA_VALUE, na.rm = TRUE),
  #               YEAR = as.numeric(YEAR))
```

##### Catch per trip
```{r}
# loop over wave
for(i in c("JAN/FEB", "MAR/APR", "MAY/JUN", "JUL/AUG", "SEP/OCT", "NOV/DEC")){
  message(i)
  
  this_data <- cpue %>%
    dplyr::filter(wave_f == i,
                  stringr::str_detect(name, "n_per_trip"),
                  !is.na(DATA_VALUE)) %>%
    dplyr::ungroup() %>%
    dplyr::group_by(st_f) %>%
    dplyr::mutate(n_years = dplyr::n()) %>%
    dplyr::filter(n_years > 2)

  this_data$name <- factor(this_data$name,
                          levels = 
                            paste(
                              c("MAINE", "NEW HAMPSHIRE", "MASSACHUSETTS",
                                     "RHODE ISLAND", "CONNECTICUT", "NEW YORK",
                                     "NEW JERSEY", "DELAWARE", "MARYLAND",
                                     "VIRGINIA", "NORTH CAROLINA", "SOUTH CAROLINA",
                                     "GEORGIA", "FLORIDA"), 
                              "\n",
                              i, 
                              "\n",
                              "n_per_trip"
                              ))
 fig <- this_data %>%
   dplyr::rename(INDICATOR_NAME = name) %>%
  esp_traffic_fig(paginate = TRUE,
                  status = FALSE,
                  silent = TRUE)
  
  print(fig)
}
```

##### Landings per trip
```{r}
# loop over wave
for(i in c("JAN/FEB", "MAR/APR", "MAY/JUN", "JUL/AUG", "SEP/OCT", "NOV/DEC")){
  message(i)
  
  this_data <- cpue %>%
    dplyr::filter(wave_f == i,
                  stringr::str_detect(name, "lbs_per_trip"),
                  !is.na(DATA_VALUE)) %>%
    dplyr::ungroup() %>%
    dplyr::group_by(st_f) %>%
    dplyr::mutate(n_years = dplyr::n()) %>%
    dplyr::filter(n_years > 2)

  this_data$name <- factor(this_data$name,
                          levels = 
                            paste(
                              c("MAINE", "NEW HAMPSHIRE", "MASSACHUSETTS",
                                     "RHODE ISLAND", "CONNECTICUT", "NEW YORK",
                                     "NEW JERSEY", "DELAWARE", "MARYLAND",
                                     "VIRGINIA", "NORTH CAROLINA", "SOUTH CAROLINA",
                                     "GEORGIA", "FLORIDA"), 
                             "\n",
                              i, 
                              "\n",
                            "lbs_per_trip"))
 fig <- this_data %>%
      dplyr::rename(INDICATOR_NAME = name) %>%
  esp_traffic_fig(paginate = TRUE,
                  status = FALSE,
                  silent = TRUE)
  
  print(fig)
}
```

## MRIP CPUE and LPUE

### Bluefish distribution

#### Catch per trip
```{r, mrippue}
wave_catch <- params$mrip_catch %>%
  dplyr::group_by(wave_f, st_f, year) %>%
  dplyr::summarise(state_catch_n = sum(tot_cat, na.rm = TRUE),
                state_land_lbs = sum(wgt_ab1, na.rm = TRUE)) %>%
  dplyr::filter(!st_f %in% c("ALABAMA", "LOUISIANA", "MISSISSIPPI")) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(year = as.numeric(year))

# directed trips by state-wave
bluefish_directed_trips <- c()
  for (i in params$trip_files) {
    this_dat <- read.csv(i,
      skip = 24,
      na.strings = "."
    )
    message(unique(this_dat$Year)) # make sure all years are downloaded
    bluefish_directed_trips <- rbind(bluefish_directed_trips, this_dat)
  }

wave_trips <- bluefish_directed_trips %>%
  dplyr::rename(wave_f = Wave, st_f = State, year = Year) %>%
    dplyr::group_by(wave_f, st_f, year) %>%
  dplyr::summarise(trips = sum(Directed.Trips, na.rm = TRUE)) %>%
  dplyr::filter(!st_f %in% c("ALABAMA", "LOUISIANA", "MISSISSIPPI")) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(wave_f = wave_f %>%
                  stringr::str_replace("JANUARY", "JAN") %>%
                  stringr::str_replace("FEBRUARY", "FEB") %>%
                  stringr::str_replace("MARCH", "MAR") %>%
                  stringr::str_replace("APRIL", "APR") %>%
                  stringr::str_replace("MAY", "MAY") %>%
                  stringr::str_replace("JUNE", "JUN") %>%
                  stringr::str_replace("JULY", "JUL") %>%
                  stringr::str_replace("AUGUST", "AUG") %>%
                  stringr::str_replace("SEPTEMBER", "SEP") %>%
                  stringr::str_replace("OCTOBER", "OCT") %>%
                  stringr::str_replace("NOVEMBER", "NOV") %>%
                  stringr::str_replace("DECEMBER", "DEC"))

cpue <- dplyr::full_join(wave_catch,
                         wave_trips,
                         by = c("wave_f", "st_f", "year")) %>%
  dplyr::mutate(n_per_trip = state_catch_n/trips,
                lbs_per_trip = state_land_lbs/trips) %>%
  dplyr::group_by(wave_f, st_f) %>%
  dplyr::summarise(mean_catch_pue = mean(n_per_trip, na.rm = TRUE),
                   mean_land_pue = mean(lbs_per_trip, na.rm = TRUE))

cpue$st_f <- factor(cpue$st_f,
                          levels = rev(c("MAINE", "NEW HAMPSHIRE", "MASSACHUSETTS",
                                     "RHODE ISLAND", "CONNECTICUT", "NEW YORK",
                                     "NEW JERSEY", "DELAWARE", "MARYLAND",
                                     "VIRGINIA", "NORTH CAROLINA", "SOUTH CAROLINA",
                                     "GEORGIA", "FLORIDA")))
cpue$wave_f <- factor(cpue$wave_f,
                            levels = c("JAN/FEB", "MAR/APR", "MAY/JUN",
                                       "JUL/AUG", "SEP/OCT", "NOV/DEC"))
cpue %>%
    dplyr::filter(wave_f != "JAN/FEB") %>%
  ggplot2::ggplot(ggplot2::aes(x = st_f,
                               y = mean_catch_pue,
                               fill = st_f)) +
  ggplot2::geom_col(color = "black") +
    ggplot2::coord_flip() +
  ggplot2::facet_wrap(~wave_f,
                      ncol = 2
           ) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::ylab("Catch (n) per trip")
```

#### Landings per trip
```{r}
cpue %>%
    dplyr::filter(wave_f != "JAN/FEB") %>%
  ggplot2::ggplot(ggplot2::aes(x = st_f,
                               y = mean_land_pue,
                               fill = st_f)) +
  ggplot2::geom_col(color = "black") +
    ggplot2::coord_flip() +
  ggplot2::facet_wrap(~wave_f,
                      ncol = 2
           ) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::ylab("Landings (lb) per trip")
```

### Annual state CPUE and LPUE
 
```{r}
cpue <- dplyr::full_join(wave_catch,
                         wave_trips,
                         by = c("wave_f", "st_f", "year")) %>%
  dplyr::group_by(st_f, year) %>%
  dplyr::summarise(annual_catch = sum(state_catch_n, na.rm = TRUE),
                   annual_land = sum(state_land_lbs, na.rm = TRUE),
                   annual_trips = sum(trips, na.rm = TRUE)) %>%
  dplyr::mutate(n_per_trip = annual_catch/annual_trips,
                lbs_per_trip = annual_land/annual_trips) %>%
  dplyr::select(-c(annual_catch, annual_land, annual_trips)) %>%
  tidyr::pivot_longer(cols = c("n_per_trip", "lbs_per_trip")) %>%
  dplyr::mutate(name = paste(st_f, name, sep = " \n ")) %>%
  dplyr::rename(YEAR = year,
                DATA_VALUE = value) %>%
  dplyr::ungroup() %>%
  # dplyr::group_by(name) %>%
  dplyr::mutate(#mean = mean(DATA_VALUE, na.rm = TRUE),
               # sd = sd (DATA_VALUE, na.rm = TRUE),
                YEAR = as.numeric(YEAR))

cpue$name <- factor(cpue$name,
                          levels = 
                            c(paste(
                              c("MAINE", "NEW HAMPSHIRE", "MASSACHUSETTS",
                                     "RHODE ISLAND", "CONNECTICUT", "NEW YORK",
                                     "NEW JERSEY", "DELAWARE", "MARYLAND",
                                     "VIRGINIA", "NORTH CAROLINA", "SOUTH CAROLINA",
                                     "GEORGIA", "FLORIDA"),
                              "\n", "n_per_trip"
                              ),
                            paste(
                              c("MAINE", "NEW HAMPSHIRE", "MASSACHUSETTS",
                                     "RHODE ISLAND", "CONNECTICUT", "NEW YORK",
                                     "NEW JERSEY", "DELAWARE", "MARYLAND",
                                     "VIRGINIA", "NORTH CAROLINA", "SOUTH CAROLINA",
                                     "GEORGIA", "FLORIDA"),
                            "\n", "lbs_per_trip")))

```

#### Catch per trip
```{r}
cpue %>%
  dplyr::filter(stringr::str_detect(name, "n_per_trip")) %>%
     dplyr::rename(INDICATOR_NAME = name) %>%
  esp_traffic_fig(paginate = TRUE,
                  status = FALSE,
                  silent = TRUE,
                  ncolumn = 2) %>% try()
```

#### Landings per trip
```{r}
cpue %>%
  dplyr::filter(stringr::str_detect(name, "lbs_per_trip")) %>%
     dplyr::rename(INDICATOR_NAME = name) %>%
  esp_traffic_fig(paginate = TRUE,
                  status = FALSE,
                  silent = TRUE,
                  ncolumn = 2)
```

